---
execute:
  cache: true
date: "2/15/2023"
author: "Jon Zelner"
---

# Re-Creating the `radon` example

First, we will load the relevant packages:
```{r setup}
#| echo: true
#| message: false
#| warning: false
library(ggplot2)
library(tidyr)
library(dplyr)
library(bayesplot)
library(rstanarm)
library(purrr)
library(tidybayes)
library(dplyr)
```


## Data Preparation

First, lets take the raw `radon` dataset from the `rstanarm` package and make some minor modifications and additonal datasets that we'll use for the purposes of modeling and visualizing these data.

First, since each row of `radon` dataset includes an observation of a single house, we need to work backwards to obtain the county-level soil uranium measure for each individual county. This is pretty straightforward to do using the `dplyr` package:

```{r}
county_uranium <- radon %>%
  group_by(county) %>%
  summarize(log_uranium = first(log_uranium)) 
```

We will also make a second dataset that we will use later on for the purposes of prediction. This contains 2 entries for each county, representing observations taken in the basement or on the first floor.

```{r}
county_uranium_tmp_1 <- county_uranium
county_uranium_tmp_1$floor <- 0
county_uranium_tmp_2 <- county_uranium
county_uranium_tmp_2$floor <- 1

county_dummy_df <- rbind(county_uranium_tmp_1, county_uranium_tmp_2)

## In the original dataset, basements are coded as the 0th floor, so we are 
## reversing this so that adjusting for basement will give us the appropriate estimate
county_dummy_df$basement <- 1-county_dummy_df$floor

## Now, lets make a final prediction dataset that includes only basement observations
pooled_predict_df <- county_dummy_df %>%
  filter(basement == 0)

## Lets also modify the `radon` dataset to have a basement variable
radon$basement <- 1 - radon$floor
```

## 🚪 Door 1: Full pooling!

```{r}
#| output: true
m1 <-lm(log_radon ~ basement, data = radon)
summary(m1)
```
Now, we'll save the fully pooled predictions for each condition to our dataset:
```{r}
county_dummy_df$pooled_pred <- predict(m1, county_dummy_df)
```
## 🚪 Door 2: No pooling
```{r}
pred_data <- data.frame(basement = c(0, 1))
no_pool_m <- lm(log_radon ~ basement + county, data = radon)
county_dummy_df$no_pool_pred <- predict(no_pool_m, county_dummy_df)
```



## 🚪 Door 3: Partial Pooling

```{r}
#| output: false
m2 <- stan_lmer(log_radon ~ basement + log_uranium + (1 | county), data = radon)
posterior <- as.matrix(m2)

## Gives posterior median for each prediction.
county_dummy_df$partial_pred <- posterior_predict(m2, county_dummy_df) %>%
  apply(2,median) 
```

```{r}
#| echo: true
g2 <- mcmc_areas(posterior, pars = c("basement", "log_uranium"))
plot(g2)
```

## Figure 1

```{r}
fig_1_counties <- c("LACQUIPARLE", "AITKIN", "KOOCHICHING", "DOUGLAS", "CLAY", "STEARNS", "RAMSEY", "STLOUIS")

county_df_fig_1 <- county_dummy_df %>% 
  filter(county %in% fig_1_counties) %>%
  mutate(county2 = factor(county, levels = fig_1_counties)) %>%
  arrange(county)

pred_counties <- radon %>% filter(county %in% fig_1_counties) %>%
  mutate(county2 = factor(county, levels = fig_1_counties))

g <- ggplot() + 
  geom_jitter(data = pred_counties, aes(x=basement, y = log_radon, group=county2),height=0, width=0.1) + 
  geom_line(data = county_df_fig_1, aes(x=basement, y=partial_pred, group=county2), linetype="solid", colour="gray") + 
  geom_line(data= county_df_fig_1, aes(x=basement, y = no_pool_pred, group = county2)) + 
    geom_line(data=county_df_fig_1, aes(x=basement, y = pooled_pred, group = county2), linetype="dashed") + 
  facet_wrap(vars(county2), nrow = 2) +
  xlab("basement") + 
  ylab("log radon level") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

plot(g)
```

## Figure 2 

```{r}
dd <- predicted_draws(m2, county_dummy_df) %>% 
  median_qi(.width=0.17) %>%
  filter(floor == 1)

uranium_coefs <- gather_draws(m2, c(`(Intercept)`, log_uranium)) %>% median_qi()

log_uranium_range <- seq(min(county_uranium$log_uranium) - .1, max(county_uranium$log_uranium) + .1, by = 0.1)

pred_log_radon <- uranium_coefs$.value[1] + uranium_coefs$.value[2]*log_uranium_range
median_radon_pred <- data.frame(log_uranium = log_uranium_range, .prediction = pred_log_radon)

g <- ggplot(dd) + 
  geom_point(aes(x = log_uranium, y = .prediction, group = county)) +
  geom_line(data = median_radon_pred, aes(x=log_uranium, y = .prediction)) + 
  geom_errorbar(aes( x= log_uranium, y= .prediction, ymin = .lower, ymax = .upper)) + 
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  xlab("county-level uranium measure") + 
  ylab("regression intercept")

plot(g)
```

