# Coming to grips with multi-level modeling

For our second zoom session, we're going to do a bit of a deep dive, hands-on take on multi-level modeling. 

## Preparation for in-class activity

In class, we will be doing a walk-through of how to reproduce the analysis in @gelman2006 using several R-based modeling tools. Before class, please:

1. **Install or update to the latest version of RStudio.** The tutorial code will be contained in a Quarto markdown document. Quarto is an updated version of the venerable RMarkdown, and the [newest versions of RStudio include Quarto support by default](https://quarto.org/docs/tools/rstudio.html). 

Before or at the beginning of class:

2. **Set up your R/RStudio installation to be able to load the following packages using the following code**:

```{r}
#| echo: true
#| output: false
library(ggplot2)
library(tidyr)
library(dplyr)
library(bayesplot)
library(rstanarm)
library(purrr)
library(tidybayes)
```

If you are not sure if you have these installed or want to update to the latest versions, please paste this command into a running R session to download and install:

```{r}
#| eval: false
install.packages(c("ggplot2","tidyr","dplyr","bayesplot","rstanarm","purrr","tidybayes"))
```

If you run into problems with any of these, please let me know!

## During Class

- [Slide](slides/session_2_multilevel) 

- You can find the tutorial we are going to work through [here](/posts/radon/index.qmd)