---
pagetitle: "Maps and more"
order: 2
format: 
    revealjs:
        theme: [solarized, styles.scss]
        incremental: true
        bibliography: /Users/jzelner/repos/bibtex-library/jz_library.bib
        cap-location: "bottom"
---

{{< include _titleslide.qmd >}}

## Agenda

- Readings

- "Good Map 😇, Bad Map 👿"

- If time: Overview of Roadmap Project

## Making sense of today's readings 

- Did the Kirby piece impact your understanding of what spatial epidemiology is or isn't? How?

- Is the distinction between spatial epidemiology and <span class="alert">medical geography</span> made by Kirby et al. meaningful?

- Which of the <span class="alert">analytic tools</span> mentioned in this piece are most relevant to your own research or practice interests?

- Any points of confusion? Other questions?

```{r}
library(countdown)
countdown(minutes = 10, warn_when = 60, seconds = 0, play_sound=FALSE)
```

## Good Map, Bad Map

- [Activity](https://zelnotes.io/teaching/EPID684/Activities/Good+Map%2C+Bad+Map) description

- [Slides](https://docs.google.com/presentation/d/1TMGl5Nl5k2K2uO-dv6VQSTEOSrPvZ3u22-7z_rInO_w/edit?usp=sharing)

```{r}
library(countdown)
countdown(minutes = 15, warn_when = 60, seconds = 0, play_sound=FALSE)
```

## Roadmap Project

- Project is really focused on how to do something useful with spatial ideas and data in public health.

- Take a few minutes to look at the components of the assignment. 

- Think about some of the ideas in the Koch piece; are they relevant to what you are interested in doing?

## Next Time

<iframe src="https://giphy.com/embed/gEfrwXsPRk14Q" width="480" height="361" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p>Making sense of distance and other forms of spatial relatedness</p>

# References
