---
author:
- Jon Zelner
header-includes:
- \usepackage{float}
- \usepackage{etaremune}
- \usepackage{wrapfig}
- \usepackage{graphicx}
- \usepackage{appendixnumberbeamer}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty, font=footnotesize}
- \usepackage{tikz}
- \usepackage{xcolor}
- \usetikzlibrary{shapes, arrows}
- \usepackage{pdfpages}

title: What is spatial epidemiology, anyway? 
institute: University of Michigan School of Public Health \newline EPID 684 \newline Winter 2022 \newline \newline `jzelner@umich.edu` \newline `epibayes.io`
fontsize: 12pt
urlcolor: blue
output: 
  beamer_presentation:
    theme: "metropolis"
    latex_engine: xelatex
    citation_package: biblatex
date: \today
bibliography: /Users/jzelner/repos/bibtex-library/jz_library.bib
---


```{r setup, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(".git/index"))

## This function uses the `here` package to make it easy to include graphics in a directory
## relative to the project root
include_local_graphics <- function(x) {
  knitr::include_graphics(here::here(x))
}
```

## Agenda

>- Who am I?
>- What is this class about? 
>- What are we going to do this semester?

## The elephant in the Zoom

```{r echo=FALSE, out.width='80%', fig.align="center"}
#| fig.cap="There is a disturbance in the...Winter term"
include_local_graphics("images/omicron_strikes_back.png")
```
../2022-01-11-11-58-20.png

## Who am I?

>- Assistant Professor of Epidemiology at UM since \alert{2017}
>- Core faculty member in the *Center for Social Epidemiology and Population Health* (CSEPH)
>- Research group focused on infectious disease ([epibayes.io](https://epibayes.io))
>- PhD in Sociology and Public Policy from UM (2011)
>- Also teach undergrad courses on the social history of infectious disease (PUBHLTH 405) 

## The final frontier?

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="Spatial analysis is closer than you might think."}
include_local_graphics("images/uss_enterprise.jpg")
```

## Course Goals

By the end of this term you will:

>1. Become effective in the application of spatial ideas to real-world problems in epidemiology.
>2. Be able to apply a spatial systems approach to public health.
>3. Learn how to select appropriate quantitative tools for addressing spatial public health problems.
>4. Understand how to use spatial analysis as a tool for advancing health justice and equity.

## Maps...but also more!

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="Disease maps have become strangely ubiquitous in the last couple of years..."}
include_local_graphics("images/nyt_covid_hotspots.png")
```


## Spatial epidemiology \alert{snuck up on me}...

```{r echo=FALSE, out.height='80%', fig.align="center", fig.cap="My first disease map, from  2015"}
include_local_graphics("images/lima_tb_incidence.png")
```

## What makes an epidemiological problem \alert{spatial}?

>- \alert{On your own:} Spend a couple of minutes collecting your thoughts about what you think makes space relevant to an epidemiological problem.

>- \alert{In small groups:}
>    - Share your ideas about spatial problems.
>    - Try to answer the following question: "What is an example of a non-spatial problem in epidemiology?"

## Spatial epidemiology is all about \alert{relationships}

>- \alert{Distance} is a stand-in for all the things that often make near things more similar and far things less similar.
>- For example: environmental exposures, social and political factors, patterns of movement, built environment.
>- Shares a lot of DNA with \alert{time-series} and \alert{network} analysis.
>- Necessarily highly interdisciplinary and open to a wide range of approaches.


## What, \alert{exactly}, are we going to be doing this term?

- [Syllabus](https://publish.obsidian.md/epibayes/Courses/EPID684/Syllabus)

## Next Time

- Digging into the [key ideas and tools of spatial epidemiology](https://publish.obsidian.md/epibayes/Courses/EPID684/modules/Module+1/Sessions/2.+Maps+and+More) in more detail.

# References
