---
fontsize: 12pt
urlcolor: blue
output: 
  revealjs::revealjs_presentation:
    theme: "solarized"
    transition: "fade"
    css: "styles.css"
    self_contained: true
    lib_dir: "libs"
    fig_caption: false
---

```{r setup, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(".git/index"))

## This function uses the `here` package to make it easy to include graphics in a directory
## relative to the project root
include_local_graphics <- function(x) {
  knitr::include_graphics(here::here(x))
}
```

# Spatial epidemiology is all about relationships {#title}

Jon Zelner   
Epid 684   
University of Michigan    
1/14/2022

# Agenda 

>- Discuss readings
>- Finish up Good Map, Bad Map activity

# The "First Law" of Geography {.quote .center}

<figure>
  <blockquote>
    "Everything is related to everything else. But near things are more related than distant things."
  </blockquote>
  <figcaption> -Waldo Tobler, 1969</figcaption>
</figure>

## {background-iframe="https://www.youtube.com/embed/kRsF9S8JqBI?autoplay=1"}

# Why is the first law useful? {.quote .center}

<figure>
<blockquote>
"[I]magine a world in which [TFL] is not true. In such a world, the full range of conditions could be encountered in every minute portion of the world. Every room, for example, might contain the full observed range of the Earth’s topographic variation, from the bottom of the Marianas Trench to the summit of Mount Everest[.] 
</blockquote>
<figcaption> -Goodchild, 2004 p. 301</figcaption>
</figure>

## "A world of white noise" {.imgslide}

```{r echo=FALSE, fig.align="center", fig.cap="TV white noise"}
include_local_graphics("images/tv_static.jpeg")
```

"White noise" has lots of variation with little autocorrelation.

# More to the story?

Goodchild argues that:

>- *Variation* is more fundamental than *autocorrelation*.
>- Without spatial variation there is only perfect autocorrelation.
>- Maybe *variation* needs to be the real First Law?
>- Doesn't discount role of correlation, just puts them in order.


# Spatial autocorrelation + variation in the real world {.imgslide .r-fit-text}

```{r echo=FALSE, out.width="60%", fig.align="center"}
#|fig.cap="Measles cases in Texarkana, Tex and Texarkana, Ark in 1970"
include_local_graphics("images/texarkana_measles.png")
```
Outbreaks of measles in two cities with very different vaccination coverages.

# {.imgslide}
<h1 class="r-fit-text">Characterizing autocorrelation and variation is tricky</h1>
```{r echo=FALSE, out.width='40%', fig.align="center", fig.cap="Checkerboard-like example of high spatial autocorrelation"}
include_local_graphics("images/morans_i_example.png")
```
Using a measure like *Moran's I* this is perfectly correlated, but we can see qualitatively that there is more going on.

# Some Questions 

Take 3m on your own:

>- Think about the epidemiological system you want to focus on for your project.
>- What kinds of mechanisms do you think are important for driving *variation* in this system?
>- How about clustering and autocorrelation?
>- Are these mechanisms more *local* or more *global*?

<p class="fragment fade-up"> Pair up and discuss for 5m.</p>

# Return of "Good Map, Bad Map"

- Activity [description](https://notes.epibayes.net/Courses/EPID684/Activities/Good+Map%2C+Bad+Map)

- [Slide deck](https://docs.google.com/presentation/d/1KJTB9Nq4S4J6aZLtjv0xTADWhEl83nuqX0KP56k8QrY/edit?usp=sharing) 

- Spend 10m finishing up slides and then come back to share.

# Next Time

- [Spatial Thinking is Systems Thinking](https://notes.epibayes.net/Courses/EPID684/modules/Module+1/Sessions/4.+Spatial+Thinking+is+Systems+Thinking)