---
author:
- Jon Zelner
header-includes:
- \usepackage{float}
- \usepackage{etaremune}
- \usepackage{wrapfig}
- \usepackage{graphicx}
- \usepackage{appendixnumberbeamer}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty, font=footnotesize}
- \usepackage{tikz}
- \usepackage{xcolor}
- \usetikzlibrary{shapes, arrows}
- \usepackage{pdfpages}
- \titlegraphic{\vspace{5cm}\flushright\includegraphics{`r here::here("assets/epibayeslogo.png")`}}

title: "EpiBayes: 2021 and Beyond"
institute: University of Michigan School of Public Health \newline `jzelner@umich.edu` \newline `epibayes.io`
fontsize: 12pt
output: 
  beamer_presentation:
    theme: "metropolis"
    latex_engine: xelatex
    citation_package: biblatex
date: \today
bibliography: /Users/jzelner/repos/bibtex-library/jz_library.bib
---


```{r setup, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(".git/index"))
## This function uses the `here` package to make it easy to include graphics in a directory
## relative to the project root
include_local_graphics <- function(x) {
  knitr::include_graphics(here::here(x))
}

```

## EpiBayes

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="For more info, check out our group site at epibayes.io"}
include_local_graphics("images/epibayes-blurb.png")
```

## Who we are:

>- **PhD students:** Kelly Broen, Ramya Naraharisetti, Hannah Steinberg, Rob Trangucci

>- **Postdoc:** Joey Dickens

>- **Research Area Specialists:** Krzysztof Sakrejda, Stephanie Choi

>- Many other faculty, postdoc, and student collaborators at UM, NIH, CDC, and beyond...

## A lot has happened in the last 18+ months!

```{r echo=FALSE, out.width='90%', fig.align="center"}
include_local_graphics("images/mi_covid_timeline_august_2021.png")
```


## What my kids looked like right before the pandemic
```{r echo=FALSE, out.width='60%', fig.align="center", fig.cap="January 2020"}
include_local_graphics("images/kids/pre-pandemic-kids.jpeg")
```

## How they look now...

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="September 2021"}
include_local_graphics("images/memes/brother-sister.jpeg")
```

## How they look now...

```{r echo=FALSE, out.width='50%', fig.align="center", fig.cap="September 2021"}
include_local_graphics("images/kids/kids-summer-2021.jpeg")
```


## Some of our \alert{big themes}

>- Inequality in infectious disease risk [@zelner2021h]

>- Spatial variation in infection outcomes [@contreras2020; @masters2020c]

>- Measurement issues in seroepidemiology [@zelner2019a; @kubale2020a]

>- Missing data & bias in estimates of key social/spatial risks 

>- Privacy and ethics in applications of spatial epidemiology [@broen2021]

## COVID has shifted what we do, but also solidified it

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="covidmapping.org, developed in collaboration with MDHHS"}
include_local_graphics("images/projects/covidmapping-august-2021.png")
```

## Explore these issues through many different \alert{disease systems}

>- COVID-19 [@zelner2020]

>- Influenza

>- Coccidiomycosis [@heaney2021]

>- Endemic Burkitt's Lymphoma

>- MRSA

>- Klebsiella Pneumoniae

# Recent and Ongoing Projects

## Inequality in \alert{COVID-19 mortality} in Michigan driven by incidence rather than case-fatality

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Disparities in COVID-19 incidence and case-fatality for March-June 2020"}
include_local_graphics("images/projects/zelner-cid-covid-disparities.png")
```

## How have these inequities been impacted by changing pandemic conditions?

```{r echo=FALSE, out.width='90%', fig.align="center"}
include_local_graphics("images/mi_covid_timeline_august_2021.png")
```

## Differences in \alert{case-fatality} have remained small over the course of the pandemic

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Figure from Naraharisetti et al., In Prep"}
include_local_graphics("images/projects/ramya-cfr-covid-2020.png")
```

## Equalizing the \alert{pacing} of infection could have prevented a large fraction of deaths among Black people in MI

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Figure from Naraharisetti et al., In Prep"}
include_local_graphics("images/projects/ramya-deaths-averted.png")
```

## \alert{Local intensity} of Malaria exposure associated with increased pediatric cancer risk

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Spatial clustering of endemic Burkitt's Lymphoma in Tanzania, Uganda and Kenya (From Broen et al., In Prep)."}
include_local_graphics("images/projects/kelly_ebl_clustering.png")
```

## Spatial clustering of pediatric non-vaccination dramatically increases VPD outbreak risk

```{r echo=FALSE, out.width='50%', fig.align="center", fig.cap="Figure from Masters et al. 2020 \\protect\\cite{masters2020}"}
include_local_graphics("images/projects/nina-pnas-clustering.png")
```

## \alert{Layered screening} and \alert{intake cohorting} are necessary to reduce COVID-19 outbreak risk in a large urban jail

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Figure from Sakrejda et al., In Prep"}
include_local_graphics("images/projects/sakrejda-ccj-screening.png")
```

## See something interesting?

- If these or other projects are of interest, let me or anyone else know!

- Don't hesitate to drop a line at `jzelner@umich.edu`

- Check out our other work at `covidmapping.org` and `epibayes.io`

- Thanks!


# References
