---
author:
- Jon Zelner
header-includes:
- \usepackage{float}
- \usepackage{etaremune}
- \usepackage{wrapfig}
- \usepackage{graphicx}
- \usepackage{appendixnumberbeamer}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty, font=footnotesize}
- \usepackage{tikz}
- \usepackage{xcolor}
- \usetikzlibrary{shapes, arrows}
- \usepackage{pdfpages}
- \titlegraphic{\vspace{5cm}\flushright\includegraphics{`r here::here("assets/epibayeslogo.png")`}}

title: "What's going on? What's next?"
institute: University of Michigan School of Public Health \newline `jzelner@umich.edu` \newline `epibayes.io`
fontsize: 12pt
output: 
  beamer_presentation:
    theme: "metropolis"
    latex_engine: xelatex
    citation_package: biblatex
date: \today
bibliography: /Users/jzelner/repos/bibtex-library/jz_library.bib
---


```{r setup, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(".git/index"))
## This function uses the `here` package to make it easy to include graphics in a directory
## relative to the project root
include_local_graphics <- function(x) {
  knitr::include_graphics(here::here(x))
}

```

## Where have we been? 

```{r echo=FALSE, out.width='60%', fig.align="center", fig.cap="What my kids looked like right before the pandemic"}
include_local_graphics("images/kids/pre-pandemic-kids.jpeg")
```

## A lot has happened!

```{r echo=FALSE, out.width='50%', fig.align="center", fig.cap="How they look now..."}
include_local_graphics("images/kids/kids-summer-2021.jpeg")
```

## A lot of \alert{growth}

Currently:

>- 4 PhD students

>- 1 postdoc

>- 2 Research Area Specialists

## Our \alert{big themes}

>- Inequality in infectious disease outcomes. [@zelner2021h]

>- Spatial heterogeneity

>- Measurement issues in seroepidemiology

>- Missing data and bias

>- Privacy, ethics and applications of spatial epidemiology [@broen2021]

## Explore these issues through many different systems

- COVID-19 [@zelner2020]

- Influenza

- Coccidiomycosis [@heaney2021]

- Endemic Burkitt's Lymphoma

- MRSA

- Klebsiella Pneumoniae


# References
