---
author:
- Jon Zelner
header-includes:
- \usepackage{float}
- \usepackage{etaremune}
- \usepackage{wrapfig}
- \usepackage{graphicx}
- \usepackage{appendixnumberbeamer}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{tikz}
- \usepackage{xcolor}
- \usetikzlibrary{shapes, arrows}
- \usepackage{pdfpages}
- \titlegraphic{\vspace{5cm}\flushright\includegraphics{`r here::here("assets/epibayeslogo.png")`}}

title: "Conceptualizing and measuring inequality in infectious disease risks"
institute: University of Michigan School of Public Health \newline `jzelner@umich.edu` \newline `epibayes.io`
fontsize: 12pt
output: 
  beamer_presentation:
    theme: "metropolis"
    latex_engine: xelatex
    citation_package: biblatex 
date: \today
bibliography: /Users/jzelner/repos/bibtex-library/jz_library.bib
---


```{r setup, echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_root(".git/index"))
## This function uses the `here` package to make it easy to include graphics in a directory
## relative to the project root
include_local_graphics <- function(x) {
  knitr::include_graphics(here::here(x))
}

```


## EpiBayes?

```{r echo=FALSE, out.width='80%', fig.align="center", fig.cap="For more info, check out our group site at epibayes.io"}
include_local_graphics("images/epibayes-blurb.png")
```

## Who we are:

>- **PhD students:** Kelly Broen, Ramya Naraharisetti, Hannah Steinberg, Rob Trangucci

>- **Postdoc:** Joey Dickens

>- **Research Area Specialists:** Krzysztof Sakrejda, Stephanie Choi

>- Many other faculty, postdoc, and student collaborators at UM, NIH, CDC, and beyond...


## COVID-19 has solidified focus on measurement and conceptualization of \alert{infection inequality} 

>- Understanding spatial variation in infection outcomes [@contreras2020; @masters2020c]

>- Missing data & bias in estimates of key social/spatial risks (Trangucci et al., In Review)

>- Privacy and data dignity in applied spatial epidemiology [@broen2021]

>- Understanding drivers of inequality infectious disease mortality [@zelner2021h], (Naraharisetti et al., In Prep)

## Inequality in \alert{early COVID-19 mortality} in Michigan driven by incidence rather than case-fatality

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Disparities in COVID-19 incidence and case-fatality for March-June 2020"}
include_local_graphics("images/projects/zelner-cid-covid-disparities.png")
```

## How were these inequities impacted by \alert{changing pandemic condition}?

```{r echo=FALSE, out.width='90%', fig.align="center"}
include_local_graphics("images/mi_covid_timeline_august_2021.png")
```

## Differences in \alert{case-fatality} have remained small over the course of the pandemic

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Figure from Naraharisetti et al., In Prep"}
include_local_graphics("images/projects/ramya-cfr-covid-2020.png")
```

## Equalizing the \alert{pacing} of infection could have prevented a large fraction of deaths among Black people in MI

```{r echo=FALSE, out.width='70%', fig.align="center", fig.cap="Figure from Naraharisetti et al., In Prep"}
include_local_graphics("images/projects/ramya-deaths-averted.png")
```

## See something interesting?

- If these or other projects are of interest, let me know.

- Don't hesitate to drop a line at `jzelner@umich.edu`

- Check out our other work at `covidmapping.org` and `epibayes.io`

- Thanks!


# References
